Which ``Bad Smells'' can be Ignored?

### 2016-03-10 Manuscript.tex
Updates from ShareLaTeX

%\usepackage{stfloats}

@@

\title{  How to Recognize  a Bad  ``Bad Smell'' using Contrast Sets}
\numberofauthors{2} 
\author{  
\alignauthor
Rahul~Krishna, Tim Menzies  \\
       \affaddr{Computer Science, NC State, USA}\\
       {\{i.m.ralk,~tim.menzies\}@gmail.com}
\alignauthor
Lucas Layman \\
       \affaddr{Fraunhofer CESE, College Park, USA}\\ 
       {llayman@cese.fraunhofer.org}
\setlength{\columnsep}{7mm}}
\maketitle
\begin{abstract} 
It is hard to assess
if a ``bad smell'' is irrelevant or misleading
for a specific project.
Developers, text books, and tools
disagree on which bad smells are important. 
Specifically, it is unclear at what threshold values
for which metrics should trigger code reorganization  since
most studies do not offer
both metrics {\em and} a threshold for their proposed bad smells.
This paper proposes the XTREE contrast set
learner to assess bad smells
and reject any that are not supported by
evidence in the historical record.  
Using XTREE, we   show that the range of useful
``bad smells'' changes from project to project. Hence,
before refactoring, it is important to run a critique tool
like XTREE to stop work on useless code reorganizations.
Further, XTREE shows that many
of the bad smells defined in standard texts and tools
are not useful for particular projects. 
Based on these results, we endorse  
guiding code reorganization via bad smells, but one must empirically validate the importance of specific bad smells within the context of each project using an approach like XTREE.

@@

On the other hand, other papers cast doubts on the value of use bad smalls
as triggers for code change.
In our own work, documented later in this paper, we  show
that tools and text books and developers can disagree on what bad smells
are important enough to hunt down and eliminate. Our findings
are consistent with several other papers.
Mantyla et al.~\cite{Mantyla2004} found that within an organization,
 developers had different perceptions on the importance of different
 bad smells.    Yamashita  and Moonen~\cite{Yamashita2013}
 surveyed 85 professional developers to conclude that it remains
 open to debate if code smells are a meaningful conceptualization  of code quality issues (from a developer's perspective). 
 Sj\"oberg et al.~\cite{Sjoberg2013} studied six developers 
 performing maintenance tasks that took three to four weeks to complete. 
 After modifying nearly 300 Java files, it was observed
 that none of the dozen investigated code smells explored in 
 that study were mostly irrelevant to to that maintenance work.

Debates on when/what code should be restructed are interesting, especially
in the current context of  widely-used SE practices such as
test-driven agile development~\cite{beck2003test,janzen05,williams2003test,george2003initial}.
In that approach, developers are always asking themselves ``should I reorganize this code''?
Bad smells are a widely used  technique for deciding when to reorganize\footnote{The Fowler text is widely read-- as of  early 2016, ACM Digital Portal reports 
it has 6434 references. Also, bad smell-based research
is widely reported-- a search  on {\em "bad smell"+programming} 
where {\em proceedings=ICSE} finds 1,748 papers since 2006.}.
Hence it is important to understand the value of
 bad smells.  
Valuable development resources can be wasted if developers
change their code based on   bad smells that are ``bad''; i.e. 
 irrelevant or misleading
for some particular project.

It turns out that it is hard to learn ``good'' bad smell detectors. Such a detector
is a combination of some code metric (e.g. lines of code in a method)
and a threshold measurement (e.g. {\em loc} $>$ 100). 
Shantnawi~\cite{Shatnawi10} warn that  there are most  studies do not offer
{\em both} a metrics and a threshold for their proposed bad smells. For example,  Darcy and Kemerer~\cite{darcy05}
discuss what OO design metrics might defect bad smells (e.g. a excessive
depth of inheritance tree) but they never comment on what thresholds
means that some section of code starts to smell bad. After
an extensive literature review on OO bad smells, Shantnawi remarks that there is a dearth
of  studies  conducted to formulate bad smell guidelines,
represented as metrics plus threshold values, that interpret the complexity of the software design using OO code metrics.

Why are there so few papers offer bad smell metrics and thresholds? For the few that
do exist, why do developers disagree on the relevance of those bad smells?
As discussed on the next page, one reason for this is that 
   cognitive biases result in different
developers declaring that
bad smell X or Y is most important.

@@

This paper proposes the XTREE contrast set learner for recognizing
``bad'' bad smells. 
Contrast set learners   recommend {\em what to change} in order to move from
     one classification to another (e.g. from defective=yes to
     defective=no).
     When applied to defect predictors from static code measures,
that delta is a change to static code measures;
i.e. a code reorganization. Hence, constrast set learners
like XTREE can be used to  recognize ``bad''
bad smells:
\begin{itemize}
    \item Consider $N$ developers each proposing a software reorganization 
based on a set of bad smells $S$.
\item XTREE will (a)~learn contrast sets
from the historical record; then (b)~reject  any reorganization based on ``bad'' bad smells, i.e.
 on any member of $S_{\mathit{bad}} \in S$
that XTREE could {\em not} be found in the historical record.
\end{itemize}
Note that we    recommend  XTREE  is used to critize, but  not to create,
bad smells. 
XTREE uses frequency counts of code features seen in the historical record\footnote{From LL: Note that ``new'' bad smells may not have a historical record. A limitation to be acknowledged and a reason why developer knowledge is still relevant.}
to generate
predictions about the effects of changing code
attributes. Such predictions may be inaccurate due to the infamous correlation-vs-causation problem
where spurious correlations can suggest effects that are not
truly causal.

@@

XTREE   has certain limits. Firstly, it is an oracle that  might reject code reorganizations proposed
by developers. XTREE does not perform those refactorings.
Secondly, XTREE assumes   Fowler's bad smells
should be assessed via their contribution to software defects
(evidence: see the section in italics of the quote
that starts this introduction). Some researchers 
 comment that   debating bad smells is  useful for more just reducing defects. For example, Bosu and Carver~\cite{bosu13} report that
debates about refactorings can   (a)~introducing new comers to the code base or (b)~share knowledge around a team
about interactions within the code. In that view, 
the concept of refactoring can be useful
even if it does not address issues of defects. 
That said, using
defect predictors,
it is possible to operationalize XTREE
while it might be harder to otherwise operationalize the assessment of bad smells.
Hence, 
this paper assumes that {\em better} bad smells and
bettering at finding code changes that  {\em reducing} the chances
that the new code will contain defects.

@@

    \item A new kind of contrast set learner called XTREE
    that   critique plans of what to change   in order
    to improve outcomes.
    \item An experimental rig for assessing the value of tools like
    XTREE that propose changes to software. That rig uses data mining
    to build an oracle for the proposed changes.
    \item Using that learner and that oracle,
    a demonstration that XTREE is better than our prior approach~\cite{me12c}
    as well as   state of the art methods~\cite{Shatnawi10,Alves2010,hermans15}
    at avoiding code reorganizations that are irrelevant to the task of reducing the risk of introducing defects.
    \end{itemize}
    Based on the above, we can offer
    a resolution between proponents of bad smells (e.g. Fowler, Shantnawi, Tufano~\cite{fowler99,Shatnawi10,Tufano2015}) and those who find them not useful or confusing   (e.g.Mantyla, Sj\"oberg,Yamahista et al.~\cite{Sjoberg2013,Mantyla2004,Yamashita2013}). We say
    that bad smells can guide code change, but only if those bad smells are
    first carefully assessed.

@@

Two of the studies in \fig{smells} offers some comments on the relative importance
of the different bad smells the   Yamashita'13 study and the developer study:
    \begin{itemize}
        \item Three of the bad smells listed in the top half of the Yamashita'13 rankings also score very high in the developer survey. Those three were {\em duplicated code, large class}, 
        and {\em long method}.
        \item
 Note that this agreement also means that the
  Yamashita'13 study and the {\em developer survey}   concluded
        that most of the  listed code smells are {\em not} high priority issues
        requiring code reorganization. 
 \end{itemize} 
In summary, just because a developer strongly believes something
about bad smells, does
not necessary mean that that belief is useful for the current project.
Human developers can be clever, but their thinking can also be distorted
by cognitive biases.
Hence, as shown in \fig{smells}, developers and text books and tools 
can disagree on which bad smells are important.
Hence it is important to have some automated tool
that can assess their beliefs in, for example,
bad smells.

@@

Secondly, the outlier approaches described above are all 
{\em univariant} which means
they do  not understand
that sets of code metrics must
be changed at the same time.
For example:
\begin{itemize} 
\item One way to remove unusually large methods is to reorganize that method's
functionality into several smaller ones. 
\item If that is done,
this means that {\em reducing} lines of code in one method
will {\em increase} the coupling of the methods of that class to
either itself or to other classes (depending on whether or not
the new smaller methods are placed in the original class, or in other classes).
\item
Hence, it can be counter-productive to act on bad smell thresholds for {\em both}
large lines of code {\em and} seemingly excessive coupling since
one cannot be reduced without increasing the other.
\end{itemize}

@@

To find an answer to this question, we will assume that developers
reorganize their code until the bad smell thresholds no longer apply.
This reorganization will start with some {\em initial} code
base that is changed to a {\em new} code base. 
For example, if the bad smell is {\em loc $>100$} and a 
code module has (e.g.) 500 lines of code, we conduct a ``what-if''
query and create a {\em new} code base with that module having {\em loc=100}.
Using a defect
predictor learned from static code metrics, we then predict the
number of defects in $d_0,d_1$ in {\em initial} and {\em new},
respectively. In this experiment,
the best bad smell threshold detector is the one that maximizes
\begin{equation}\label{eq:diff}
\mathit{improvement} = 100* \left(1 - \frac{ d_1 }{ d_0}\right)
\end{equation}

@@

\subsection{Resylts}

>>>>

\usepackage{rotating}
%\usepackage{stfloats}

@@

\title{  Which ``Bad Smells'' can be Ignored?}
\numberofauthors{2} 
\author{  
\alignauthor
Rahul~Krishna, Tim Menzies  \\
       \affaddr{Computer Science, NC State, USA}\\
       {\{i.m.ralk,~tim.menzies\}@gmail.com}
\alignauthor
Lucas Layman \\
       \affaddr{Fraunhofer CESE, College Park, USA}\\ 
       {llayman@cese.fraunhofer.org}
\setlength{\columnsep}{7mm}}
\maketitle
\begin{abstract} 
The literature lists dozens of bad smells which might
be detrimental to a software system.
But which are relevant to the current project
and which can be ignored?
Developers, text books, and tools
disagree on which bad smells are important. 
Specifically, it is unclear at what threshold values
a measurement should trigger code refactorings  since
most studies do not offer
{\em both} measurable metrics {\em and} a threshold for their proposed bad smells.

This paper proposes   XTREE, a tool to  prioritize bad smells. Given an historical
log of defects
seen previously in the code, XTREE
  deprioritizes   bad smells  that have little to no support
  in that log.

  This paper compares the bad smells found by XTREE to
  those found by previous work by Shatnawi, Alves, and Borges. 
  If code modules are  changed to avoid the bad
  smells found by XTREE,  that changed system
  has significantly fewer expected defects.
  Also, XTREE   endorses
  changes to far fewer code metrics.
  Hence, it is 
  a better tool for ruling out irrelvant bad smells.
  For example, the XTREE results show that many of the bad smells learned 
  by Shatnawi, Alves, and Borges can be ignored.

Based on these results, we endorse using bad smells to  
guiding refactoring. However, not all bad smells
are   relevant to the current project.
Hence, it is important
to  prioritize the importance of specific bad smells within the context of each project.

@@

On the other hand, other papers cast doubts on the value of use bad smalls
as triggers for code change.
As we show in this paper,  
 tools and text books and developers can disagree on what bad smells
are important enough to hunt down and eliminate. Our findings
are consistent with several other papers.
Mantyla et al.~\cite{Mantyla2004} found that within an organization,
 developers had different perceptions on the importance of different
 bad smells.    Yamashita  and Moonen~\cite{Yamashita2013}
 surveyed 85 professional developers to conclude that it remains
 open to debate if code smells are a meaningful conceptualization  of code quality issues (from a developer's perspective). 
 Sj\"oberg et al.~\cite{Sjoberg2013} studied six developers 
 performing maintenance tasks that took three to four weeks to complete. 
 After modifying nearly 300 Java files, it was observed
 that none of the dozen investigated code smells explored in 
 that study were mostly irrelevant to to that maintenance work.

Debates on when/what code should be restructed are interesting, especially
in the current context of  widely-used SE practices such as
test-driven agile development~\cite{beck2003test,janzen05,williams2003test,george2003initial}.
In that approach, developers are always asking themselves ``should I refactor this code''?
Bad smells are a widely used  technique for deciding when to refactor\footnote{Evidence: As of  early 2016, ACM Digital Portal reports 
the Fowler text has 6434 references. Also, bad smell-based research
is widely reported. A search  on {\em "bad smell"+programming} 
where {\em proceedings=ICSE} finds 1,748 papers since 2006.}.
Hence it is important to understand the value of
 bad smells.  
Valuable development resources can be wasted if developers
refactor their code based on   bad smells that are ``bad''; i.e. 
 irrelevant or misleading
for some particular project.

It turns out that it is hard to learn ``good'' bad smell detectors.
As discussed on the next page, programmer
   cognitive biases result in different
developers declaring that
bad smell X or Y is most important.

Also, a bad smell  detector
is a combination of some code metric (e.g. lines of code in a method)
and a threshold measurement (e.g. {\em loc} $>$ 100). 
Shantnawi~\cite{Shatnawi10} warn that  there are most  studies do not offer
{\em both} a metrics and a threshold for their proposed bad smells. For example,  Darcy and Kemerer~\cite{darcy05}
discuss what OO design metrics might defect bad smells (e.g. a excessive
depth of inheritance tree) but they never comment on what thresholds
means that some section of code starts to smell bad. After
an extensive literature review on OO bad smells, Shantnawi remarks that there is a dearth
of  studies  conducted to formulate bad smell guidelines,
represented as metrics plus threshold values, that interpret the complexity of the software design using OO code metrics.

Further, it is hard to judge the  effects of removing bad smells
for three reasons.
Firstly, the bad smell learners currently found in the literature
report bad smells for each code metric {\em in isolation}
Refactored code cannot be assessed just by a rerun of the the test
suite since
code refactorings is meant to
preserve the semantics of the code (i.e. the same tests
should pass before and after the refactoring). 
It is also problematic to assessing refactorings by making the
refactorings,  then seeing what happens to that code for the rest
of the project. In an test-driven agile project, as
SCRUM meetings change what stories are implemented next,
the operational profile that lead to defects in the old
code may not repeat.

To address the first problem of programmer cognitive biases,
this paper proposes using the   XTREE contrast set learner to
find bad smells. Contrast set learners   recommend {\em what to change} in order to move from
     one classification to another (e.g. from defective=yes to
     defective=no).
     When applied to defect predictors from static code measures,
that delta is a change to static code measures;
i.e. a code refactoring. Given an historical log of defects seen previously in the code, XTREE deprioritizes bad smells that have little to no support in that log.

Note that XTREE also addresses the second problem since it reports
code metrics and useful thresholds that

To resolve this third problem, this paper adopts
a technique used by Harman et al.

Why are there so few papers offer bad smell metrics and thresholds? For the few that
do exist, why do developers disagree on the relevance of those bad smells?

@@

This paper proposes the XTREE contrast set learner for recognizing
``bad'' bad smells. 
Contrast set learners   recommend {\em what to change} in order to move from
     one classification to another (e.g. from defective=yes to
     defective=no).
     When applied to defect predictors from static code measures,
that delta is a change to static code measures;
i.e. a code refactoring. Hence, constrast set learners
like XTREE can be used to  recognize ``bad''
bad smells:
\begin{itemize}
    \item Consider $N$ developers each proposing a software refactoring 
based on a set of bad smells $S$.
\item XTREE will (a)~learn contrast sets
from the historical record; then (b)~reject  any refactoring based on ``bad'' bad smells, i.e.
 on any member of $S_{\mathit{bad}} \in S$
that XTREE could {\em not} be found in the historical record.
\end{itemize}
Note that we    recommend  XTREE  is used to critize, but  not to create,
bad smells. 
XTREE uses frequency counts of code features seen in the historical record\footnote{From LL: Note that ``new'' bad smells may not have a historical record. A limitation to be acknowledged and a reason why developer knowledge is still relevant.}
to generate
predictions about the effects of changing code
attributes. Such predictions may be inaccurate due to the infamous correlation-vs-causation problem
where spurious correlations can suggest effects that are not
truly causal.

@@

XTREE   has certain limits. Firstly, it is an oracle that  might reject   refactorings proposed
by developers. XTREE does not perform those refactorings.
Secondly, XTREE assumes   Fowler's bad smells
should be assessed via their contribution to software defects
(evidence: see the section in italics of the quote
that starts this introduction). Some researchers 
 comment that   debating bad smells is  useful for more just reducing defects. For example, Bosu and Carver~\cite{bosu13} report that
debates about refactorings can   (a)~introducing new comers to the code base or (b)~share knowledge around a team
about interactions within the code. In that view, 
the concept of refactoring can be useful
even if it does not address issues of defects. 
That said, using
defect predictors,
it is possible to operationalize XTREE
while it might be harder to otherwise operationalize the assessment of bad smells.
Hence, 
this paper assumes that {\em better} bad smells and
bettering at finding code changes that  {\em reducing} the chances
that the new code will contain defects.

@@

    \item A new kind of contrast set learner called XTREE
    that   critique plans of what to change   in order
    to improve outcomes.
    \item An experimental rig for assessing the value of tools like
    XTREE that propose changes to software. That rig uses data mining
    to build an oracle for the proposed changes.
    \item Using that learner and that oracle,
    a demonstration that XTREE is better than our prior approach~\cite{me12c}
    as well as   state of the art methods~\cite{Shatnawi10,Alves2010,hermans15}
    at avoiding refactorings that are irrelevant to the task of reducing the risk of introducing defects.
    \end{itemize}
    Based on the above, we can offer
    a resolution between proponents of bad smells (e.g. Fowler, Shantnawi, Tufano~\cite{fowler99,Shatnawi10,Tufano2015}) and those who find them not useful or confusing   (e.g.Mantyla, Sj\"oberg,Yamahista et al.~\cite{Sjoberg2013,Mantyla2004,Yamashita2013}). We say
    that bad smells can guide code change, but only if those bad smells are
    first carefully assessed.

@@

Two of the studies in \fig{smells} offers some comments on the relative importance
of the different bad smells the   Yamashita'13 study and the developer study:
    \begin{itemize}
        \item Three of the bad smells listed in the top half of the Yamashita'13 rankings also score very high in the developer survey. Those three were {\em duplicated code, large class}, 
        and {\em long method}.
        \item
 Note that this agreement also means that the
  Yamashita'13 study and the {\em developer survey}   concluded
        that most of the  listed code smells are {\em not} high priority issues
        requiring refactoring. 
 \end{itemize} 
In summary, just because a developer strongly believes something
about bad smells, does
not necessary mean that that belief is useful for the current project.
Human developers can be clever, but their thinking can also be distorted
by cognitive biases.
Hence, as shown in \fig{smells}, developers and text books and tools 
can disagree on which bad smells are important.
Hence it is important to have some automated tool
that can assess their beliefs in, for example,
bad smells.

@@

Secondly, the outlier approaches described above are all 
{\em univariant} which means
they do  not understand
that sets of code metrics must
be changed at the same time.
For example:
\begin{itemize} 
\item One way to remove unusually large methods is to refactor that method's
functionality into several smaller ones. 
\item If that is done,
this means that {\em reducing} lines of code in one method
will {\em increase} the coupling of the methods of that class to
either itself or to other classes (depending on whether or not
the new smaller methods are placed in the original class, or in other classes).
\item
Hence, it can be counter-productive to act on bad smell thresholds for {\em both}
large lines of code {\em and} seemingly excessive coupling since
one cannot be reduced without increasing the other.
\end{itemize}

@@

To find an answer to this question, we will assume that developers
refactor their code until the bad smell thresholds no longer apply.
This refactor will start with some {\em initial} code
base that is changed to a {\em new} code base. 
For example, if the bad smell is {\em loc $>100$} and a 
code module has (e.g.) 500 lines of code, we conduct a ``what-if''
query and create a {\em new} code base with that module having {\em loc=100}.
Using a defect
predictor learned from static code metrics, we then predict the
number of defects in $d_0,d_1$ in {\em initial} and {\em new},
respectively. In this experiment,
the best bad smell threshold detector is the one that maximizes
\begin{equation}\label{eq:diff}
\mathit{improvement} = 100* \left(1 - \frac{ d_1 }{ d_0}\right)
\end{equation}

@@

\subsection{Results}
\input{tex/results_jur.tex}
% \input{tex/results_shatnawi.tex}
% \input{tex/results_alves.tex}

### 2016-03-10 tex/refactor.tex
Updates from ShareLaTeX

\begin{figure*}[!t] 
\scriptsize
\centering
\begin{tabular}{r|c|c|c|c|c}
Mostly from Fowler'99~\cite{fowler99}& & & & & Developer\\
 (with a some additions)~\cite{fowler99} & Lanza'06~\cite{Lanza2006} & SonarQube~\cite{sq15}  &  Yamashita'13\cite{Yamashita2013} & Atwood'06\cite{Atwood06}  & Survey 2015\\\hline
  Alt. Classes with Diff. Interfaces & & & & \checkmark & \\
  Combinatorial Explosion~\cite{Kerievsky2005} & & & & \checkmark & \\
  Comments & & & 11 & \checkmark & VL\\
  Conditional Complexity~\cite{Kerievsky2005} & & & 14 & \checkmark & ?\\
  Data Class & \checkmark & & & \checkmark &\\
  Data Clumps &  &  & & \checkmark &\\
  Divergent Change & & & & \checkmark & \\
  Duplicated Code & \checkmark & \checkmark & 1 & \checkmark & VH\\
  Feature Envy & \checkmark & & 8 & \checkmark &\\

@@

\caption{List of bad code smells from different sources.  Check marks (\checkmark) denote that a bad smell was mentioned.
Numbers of symbolic labels (e.g. "VH") denote  a priorization comment.
``?'' indicates lack of consensus on some bad smell. Empty cells
denote some bad smell listed in column one that was not found relevant
in subsequent studies.
Note that there are many blank cells.}
\label{fig:smells}
\end{figure*}

>>>>

\begin{figure}[!t] 
\scriptsize
\centering
\begin{tabular}{r|c|c|c|c|c}

\begin{turn}{75}Fowler'99~\cite{fowler99}, and~\cite{Kerievsky2005})\end{turn} &\begin{turn}{75} Lanza'06~\cite{Lanza2006}\end{turn} & \begin{turn}{75}SonarQube~\cite{sq15} \end{turn} &  \begin{turn}{75}Yamashita'13\cite{Yamashita2013} \end{turn}& \begin{turn}{75}Atwood'06\cite{Atwood06}\end{turn}  & \begin{turn}{75} Developer Survey 2015\end{turn}\\\hline
  Alt. Classes with Diff. Interfaces & & & & \checkmark & \\
  Combinatorial Explosion~\cite{Kerievsky2005} & & & & \checkmark & \\
  Comments & & & 11 & \checkmark & VL\\
  Conditional Complexity~\cite{Kerievsky2005} & & & 14 & \checkmark & ?\\
  Data Class & \checkmark & & & \checkmark &\\
  Data Clumps &  &  & & \checkmark &\\
  Divergent Change & & & & \checkmark & \\
  Duplicated Code & \checkmark & \checkmark & 1 & \checkmark & VH\\
  Feature Envy & \checkmark & & 8 & \checkmark &\\

@@

\caption{List of bad code smells from different sources.  Check marks (\checkmark) denote that a bad smell was mentioned.
Numbers of symbolic labels (e.g. "VH") denote  a priorization comment.
``?'' indicates lack of consensus on some bad smell. Empty cells
denote some bad smell listed in column one that was not found relevant
in subsequent studies.
Note that there are many blank cells.}
\label{fig:smells}
\end{figure}
